Global code motion of congruent instructions

### 2017-04-18 paper.tex
some more rewording

% Explain the overall algorithm, explain how anticipability is computed on demand
\begin{verbatim}
int doCodeHoisting()
  For each value number VN with 2 or more instructions:
    sort the instructions according to DFSIn numbers
    if first two I1, I2 are hoistable:
      if they are downward-safe at hoist-point:
        if they are profitable to be hoisted at hoist-point:
          proceed to check hoistability of subsequent instruction with same VN
        else
          proceed to check if I2 and I3 (if available) is hoistable

  For each hoistable VN:
    move the first one I1 to the hoist-point
    update the use of all other candidates to refer to I1
    remove all others.
    if I1 has memory references (a load, store etc.)
      update MemorySSA of I1 and others to point to the MemorySSA reference of I1
      remove MemorySSA reference of all others which were deleted
    update statistics

@@

\section{Conclusion and Future Work}
\label{sec:future-work}
We have presented the \gcm{} of congruent computations in SSA
form. We saw that it improves inlining and LICM opportunities, reduces register
spills, it also improves performance in some cases. To preserve performance and
not hoist too much we have register pressure aware cost models described in
Section~\ref{subsec:cost-models}. A part of this implementation is already merged
in LLVM trunk as GVNHoist.cpp. We will post the \GCM{} implementation for review.

There are a few improvements that we would like to make. First would be to
integrate this with \LLVM{'}s new GVN-PRE implementation such that more
congruent instructions can be scheduled. This would also expose more
redundancies to the GVN-PRE.  In general it is a good idea to start with lower
ranked expression first such that maximum hoisting can happen in one iteration,
however, current implementation does not rank the expressions and iteratively
finds a fixed point when no more candidates are available. Even this
implementation converges quickly and no compile time regression have been
observed because of code hoisting pass. This is not the most optimal approach
and results in multiple data structures to be recomputed. This can be improved
by ranking the computations \cite{rosen1988global}. Also, since \GCM{} runs very
early in the pass pipeline, it will be good to evaluate the codesize/performance
impact when it is run in sync with \GVN{}-\PRE{} just like \GCC{} does.

With the implementation of \GCM{} in \LLVM{}, the passes which rely on the
code-size/instruction-count to make optimization decisions needs to be
revisited. The first candidate would be the inliner. We have seen different
inlining decisions in Table~\ref{tab:code-size}, before and after code-hoisting
was enabled.  Since inliner has several magic numbers tuned for the previous
pass layout, it would need some improvement.

>>>>

% Explain the overall algorithm, explain how anticipability is computed on demand
\begin{verbatim}
int doCodeHoisting()
  For each value number VN with 2 or more instructions:
    sort the instructions according to DFSIn numbers
    if first two I1, I2 are hoistable:
      if they are downward-safe at hoist-point:
        if they are profitable to be hoisted at hoist-point:
          proceed to check ability to hoist of subsequent instruction with same VN
        else
          proceed to check if I2 and I3 (if available) can be hoisted

  For each set of instructions (I1, ..., Ik) with same VN that can be hoisted:
    move the first one I1 to the hoist-point
    update the use of all other candidates (I2, ..., Ik) to refer to I1
    remove all other instructions (I2, ..., Ik)
    if I1 has memory references (a load, store etc.)
      update MemorySSA of I1 and others to point to the MemorySSA reference of I1
      remove MemorySSA reference of all others which were deleted
    update statistics

@@

\section{Conclusion and Future Work}
\label{sec:future-work}
We have presented the \gcm{} of congruent computations in SSA form. We saw that
it improves inlining and LICM opportunities, reduces register spills, it also
improves performance in some cases. To preserve performance and not hoist too
much we have presented some cost models to lower register pressure. A part of
this implementation is already merged in LLVM trunk as GVNHoist.cpp, and we
posted the \GCM{} implementation for review.

There are a few improvements that we would like to make. First would be to
integrate this with \LLVM{'}s new GVN-PRE implementation such that more
congruent instructions can be scheduled. This would also expose more
redundancies to the GVN-PRE.  In general it is a good idea to start with lower
ranked expression first such that maximum hoisting can happen in one iteration,
however, current implementation does not rank the expressions and iteratively
finds a fixed point when no more candidates are available. Even this
implementation converges quickly and no compile time regressions have been
observed because of code hoisting pass. This is not the most optimal approach
and results in multiple data structures to be recomputed. This can be improved
by ranking the computations \cite{rosen1988global}. Also, since \GCM{} runs very
early in the pass pipeline, it will be good to evaluate the codesize/performance
impact when it is run in sync with \GVN{}-\PRE{} just like \GCC{} does.

With the implementation of \GCM{} in \LLVM{}, the passes which rely on the
code-size/instruction-count to make optimization decisions need to be
revisited. The first candidate would be the inliner. We have seen different
inlining decisions in Table~\ref{tab:code-size}, before and after code-hoisting
was enabled.  Since inliner has several magic numbers tuned for the previous
pass layout, it would need some improvement.
